---
title: "Viz 2"
output: github_document
---

```{r setup , include = FALSE}
library(tidyverse)
library(ggridges)
devtools::install_github("p8105/p8105.datasets")

library(p8105.datasets)
#library(readxl)
#library(haven)
```

```{r}
library(p8105.datasets)
data("weather_df")
```

## Scales

```{r}
ggp_weather_scatterplot = 
  weather_df |>
ggplot( aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = 0.5)+
  labs (
    x = "minimum daily temp",
    y = "maximum daily temp",
    title = "Temperature scatterplot",
    caption = "Data from NOAA",
    color = "Location"
  ) + #label the x axis
  scale_x_continuous(
    breaks = c (-20, 0,25),
    labels = c ("-20C", "0", "25")
  ) +
  scale_y_continuous(
    trans = "sqrt"
    #limits = c (10,30) # to zoom in, or can do filter command from the beginning 
  ) +
  #scale_color_hue(h = c(100,300))
  viridis::scale_color_viridis(
    discrete = TRUE
  )
ggp_weather_scatterplot
```
We can use the viridis color pallet

## Theme
```{r}
ggp_temperature = 
  weather_df |>
ggplot( aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = 0.5)+
  labs (
    x = "minimum daily temp",
    y = "maximum daily temp",
    title = "Temperature scatterplot",
    caption = "Data from NOAA",
    color = "Location"
  ) + #label the x axis
  scale_x_continuous(
    breaks = c (-20, 0,25),
    labels = c ("-20C", "0", "25")
  ) +
  scale_y_continuous(
    trans = "sqrt"
    #limits = c (10,30) # to zoom in, or can do filter command from the beginning 
  ) +
  #scale_color_hue(h = c(100,300))
  viridis::scale_color_viridis(
    discrete = TRUE
  )
```

Update 
```{r}
ggp_temperature +
  theme_minimal()+
  theme(legend.position = "bottom")
```

## Adding data in geoms

```{r}
central_park_df =
  weather_df |>
  filter(name == "CentralPark_NY")
molokai_df = 
  weather_df |>
  filter (name == "Molokai_HI")

ggplot (data = molokai_df, aes (x = date, y = tmax, color = name)) + 
  geom_point() +
  geom_line (data = central_park_df)
```

 ## `patchwork`
 Make three plots
```{r}
ggp_tmax_tmin = 
  weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5)
ggp_tmax_tmin
ggp_prec_density = 
  weather_df |>
  filter (prcp > 20)|>
  ggplot(aes(x = prcp, fill = name)) + 
  geom_density(alpha = 0.5)
ggp_prec_density

```

## Data manipulation
```{r}
weather_df |>
  mutate (name = fct_relevel(name, c ("Molokai_HI", "CentralPark_NY", "Waterhole_WA"))) |>
  ggplot (aes(x = name, y = tmax, fill = name)) + 
  geom_violin(alpha = 0.5)
```
 Can arrange in order to show clear data trend
```{r}
weather_df |>
  mutate ( name = fct_reorder(name, tmax)) |>
  ggplot (aes(x = name, y = tmax, fill = name)) + 
  geom_violin(alpha = 0.5)
```
 What about data tidiness
```{r}
pulse_df = 
  haven:: read_sas("./data/public_pulse_data.sas7bdat")|>
  janitor::clean_names() |>
  pivot_longer(
    bdi_score_bl : bdi_score_12m,
    names_to = "visit",
    values_to = "bdi"
  )|>
  mutate(visit = fct_inorder(visit))

```
 
 ## How to think of what you can plot: sketch the plot
 
 what is in the x : dose
 Y is the days of follow up
 facet : a column of day of treatment
 facet will be outcomes
 Write out your dataframe so you can make the plot
 
 Make a plot for FAS study
```{r}

pups_df = 
  read_csv ("./data/FAS_pups.csv", na = c("NA", ".", ""), skip = 3) |>
  janitor::clean_names()|>
  mutate (
    sex = case_match(
      sex, 
      1 ~ "male",
      2 ~ "female"
    )
  )
litter_data = 
  read_csv("./data/FAS_litters.csv", na = c("NA", ".", "")) |>
  janitor::clean_names() |>
  separate(group, into = c("dose", "day_of_tx"), sep = 3)

fas_data = left_join(pups_df, litter_data, by = "litter_number") 

fas_data |> 
  select(sex, dose, day_of_tx, pd_ears:pd_walk) |> 
  pivot_longer(
    pd_ears:pd_walk,
    names_to = "outcome", 
    values_to = "pn_day") |> 
  drop_na() |> 
  mutate(outcome = forcats::fct_reorder(outcome, pn_day, median)) |> 
  ggplot(aes(x = dose, y = pn_day)) + 
  geom_violin() + 
  facet_grid(day_of_tx ~ outcome)
```
 
